Modelling last glacial cycle ice dynamics in the Alps

### 2017-09-13 articles/cycle/alps-cycle.tex
Draft trimlines discussion.

    \begin{itemize}
    \item Ice surface is well above the trimlines (Fig.~\ref{fig:lgmvel}a).
    \end{itemize}

@@

    \begin{figure}
      \centerline{\includegraphics{alpcyc_hr_trimlines}}
      \caption{%
        \textbf{(a)} Comparison of modelled ice surface elevation at the LGM
        (time-transgressive, corresponding to maximum ice thickness,
        Fig.~\ref{fig:timing}), compensated for bedrock deformation, against
        observed trimline elevations \citep[Table~1]{Kelly.etal.2004}. Model
        variables were bilinearly interpolated to the trimline locations.
        \textbf{(b)} Histogram of differences between modelled LGM ice surface
        elevation and trimline elevations (500\,m bands). The average
        difference is 861\,m.}
        \textbf{(c)} Modelled age of maximum ice thickness at the trimline
        locations \citep[Table~1]{Kelly.etal.2004} in the upper Rhone valley
        (colour) and LGM ice surface elevation (200\,m contours). Hatches
        mark areas modelled to have experienced less than 1\,ka temperate ice
        cover during MIS~2 (29--14\,ka).
      \label{fig:trimlines}
    \end{figure}

>>>>

    In several places across the Alps, the maximum elevation reached by the
    ice surface during the LGM has been reconstructed from trimlines, the upper
    limit of glacial erosion observed in the landscape. This include the entire
    western Alps, the Rhone Glacier catchment, the Rhine Glacier catchment,
    south-eastern Switzerland and all of Austria. These reconstructions are
    based on the assumption that glacial ice was significantly erosive up to
    the or almost up to the surface, which was then located at or slightly
    above the elevation of the trimlines. This assumption is supported by the
    fact that there exist a regional consistence in trimlines elevations, which
    are generally located along a surface sloping in the assumed direction of
    flow. However, in other glaciated mountain regions, including several
    sites in the Scandinavian Mountains, the Cairgorns in Scotland, and high
    plateaus in the Canadian Cordillera, that cold-based ice commonly occured
    at high elevation and preserved the pre-glacial landscape.

    In the upper Rhone valley, the maximum ice surface elevation reached
    during MIS~2 in our simulation, corrected for bedrock deformation, is
    consistently modelled to have occured several hundred metres above the
    observed trimline elevations (Fig.~\ref{fig:trimlines}a, b), with a mean
    difference of 861\,m (xx\,\%, Fig.~\ref{fig:trimlines}b). This result is
    somewhat depedent on uncertain basal sliding and ice rheological parameters
    to which the model sensitivity was not tested here. However, a previous
    sensitivity study conducted with PISM and a very similar model set-up on
    the Cordillera ice sheet show that varying ice rheological parameters
    resulted relative errors of 25\,\unit{\%} with regard to ice rheological
    parameters and 14\,\unit{\%} regarding basal sliding parameters
    \citep[Fig.~7]{Seguinot.etal.2016}.

    In addition, our model results depict a polythermal LGM Alpine ice cover.
    Due to sub-freezing temperatures applied in the climate forcing of the
    model, the entire Alpine ice field is capped by an upper layer of cold ice.
    In major Alpine valleys, such as the upper Rhone Valley, important ice
    thickness and strain heating contribute to form a layer of temperate ice
    near the glacier base, allowing basal melt, sliding and potentially
    erosion. On the other hand, on the highest mountains ice cover is to thin
    and too static to form temperate ice, resulting in cold ice down to the
    bed, preventing potential erosion.

    In the upper Rhone valley, the geographical location of observed trimlines
    show remarkable agreement with the areal extent of regions that are
    modelled to have experienced warm-based for less than 1\,ka (7\,\%) of the
    MIS~2. Although this result calls for more detailed comparisons across the
    entire Alpine range and specific sensitivity studies to relevant basal
    sliding and ice rheological parameters, the presence of an upper layer of
    cold ice during the LGM, already found in some places in a much warmer
    climate today, is inevitable. Thus we want to challenge the assumption that
    Alpine trimlines mark the maximum upper ice surface elevation of the LGM
    ice cover and call for a more accurate estimation of the thickness of the
    upper layer cold ice in the Alps.

    \todo{look for references.}

@@

    \begin{figure}
      \centerline{\includegraphics{alpcyc_hr_trimlines}}
      \caption{%
        \textbf{(a)} Comparison of modelled ice surface elevation at the LGM
        (time-transgressive, corresponding to maximum ice thickness,
        Fig.~\ref{fig:timing}), compensated for bedrock deformation, against
        observed trimline elevations \citep[Table~1]{Kelly.etal.2004}. Model
        variables were bilinearly interpolated to the trimline locations.
        \textbf{(b)} Histogram of differences between modelled LGM ice surface
        elevation and trimline elevations (500\,m bands). The average
        difference is 861\,m.
        \textbf{(c)} Modelled age of maximum ice thickness at the trimline
        locations \citep[Table~1]{Kelly.etal.2004} in the upper Rhone valley
        (colour) and LGM ice surface elevation (200\,m contours). Hatches
        mark areas modelled to have experienced less than 1\,ka temperate ice
        cover during MIS~2 (29--14\,ka).
        \todo{add relative mean difference.}}
      \label{fig:trimlines}
    \end{figure}
